---
title: "Import Blot Data"
author: "Douglas A. Campbell"
date: "`r format(Sys.Date())`"
#output: html_notebook
---
Changes for Brian to implement:
- Fcp6SpeciesPlot aMean fmol_ugChlA (do for all all targets) (confirm if up to date)
- look through notes
- shiny interactive plot
  - species
  - blot id
  - Dany excel tab
- change plate read - unknown if not number (don't start with P)

This .Rmd RNotebook imports a metadata Catalog  with information on samples.

It imports blot data in .csv form saved in a folder after export from BioRad software based upon project specific values for variables set by the user.

It tidies and organizes the data.
It merges the metadata with the imported data based upon shared values for the sample code variable (ex. MeBe1639) which should unambiguously identify a given sample
It plots Adj_Vol_Int vs. fmol standard lane-1 for each saved blot file.
It fits a linear regression and an exponential regression to the standard curve data from each file.
It generates quantitations on the basis of the linear and exponential regressions.
It generates flag columns exp_flag & lin_flag to identify rows where the exp or lin estimates of fmol_lane fall within (1) or outside (0) the  range of the standard curve used in that filename (blot). The range was extended by 10% to increase the dynamic range of the quantitation limits.
It exports a summary of the blot quantitation data as .csv & as .Rds

It merges data with sample prep data on [protein] and load per well for each target protein.

To do:

Plot results vs. well and vs. sample category for data exploration
(Possibly a separate workbook?)

-add polynomial or logistic models for standard curves?
-put in a fit quality selection criteria.
-put in user defined standard trimming ranges

-Export the Figures

Older shut off now:
Import blot data already standardized vs. standard protein loads in Excel.


```{r set project variables}
Project <- "GreenEdge"
DataIn <- "RawData"
PlotsPath <- "Plots"
DataOut <- "ProcessData"
ProteinTargets <- c("PsbA", "RbcL", "FtsH", "Fcp6", "Lhcx6")
CatalogSheets <- c("Nf_Growth vs. Ice Camp", "LightStress_IceCamp_mimic", "Cn_Sp_pannel")
TotalProtein <- "TotalProtein"
#Fcp6 is standardized using a pseudo Phaedactyum t. std. units = ug not fmole.  This workbook assumes unit is fmole for this target.

FileEncode <- "UTF-8" 
Delimiter <- ","
#this is brittle; better to figure out how to read all data starting at line that contains a key character string
HeaderRows <- 1

#FlagRange is 10% of standard range for flagging estimates outside standard range.
FlagRange <- 0.1

```

Install Softmax XML reader package (uncomment and run once).
```{r install plate read}
# if(!require("devtools")) install.packages("devtools")
# devtools::install_github("briandconnelly/softermax", build_vignettes = TRUE)
```


```{r load libraries} 
# libraries; Note check actual dependencies
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(knitr)
library(minpack.lm)
library(nlstools)
library(nlsMicrobio)
library(readxl)
library(softermax)
library(modelr)
```

```{r load metadata Catalog and sample prep}
# initialize catalog to be appended to
Catalog <- NULL

# loop through excel sheets
for(sheet in CatalogSheets){
 Data <- readxl::read_xlsx("Proteins_Samples_update_Dany_Johann.xlsx", sheet = sheet) %>%
   # convert all column types to char for bind_rows
     mutate_all(as.character)
 # append data from the current sheet to a t
 Catalog <- bind_rows(Catalog, Data)
}

# read sample prep data
SampPrep <- as_tibble(read_excel(file.path("GreenEdgeTargetProteinSamplePrep.xlsx", fsep = .Platform$file.sep)))

#filter header rows and filter rows of all na
SampPrep <- SampPrep %>%
  slice(-1,-2) %>%
  filter_all(any_vars(!is.na(.))) %>%
  # remove manually entered protein data
  select(-Rep1TotalProteinExtract_ug_uL, 
         -Rep2TotalProteinExtract_ug_uL,
         -Rep3TotalProteinExtract_ug_uL,
         -AverageTotalProteinExtract_ug_uL,
         -TotalProteinExtractFileName)
```

```{r tidy metadata}
#Figure out how to pass value of catname GreenEdge_Catalog into following tidy pipe
Catalog <- Catalog %>%
  # re-parse all column types
  type_convert() %>%
  # date column is parsed automatically, no need to manually specify
  rename(CultureId = `Culture ID`,
         Growth_uE = `E growth (UML)`,
         Exp_uE = `E exp (UML)`,
         Stress_h = `t stress (h)`,
         DarkPre_h = `Dark pre-exp (h)`,
         Harvest_mL = `Vol (ml)`,
         Chla_ug_L = `Chl a µg/L`,
         Carbon_ug_mL = `Carbon µg/mL`,
         TotalCarbon_ug = `Total Carbon (µg)`,
         TotalChla_ug = `Total Chl a (µg)`,
         N_ug_mL =`Nitrogen µg/mL`,
         TotalN_ug = `Total nitrogen (µg)`,
         TotalProteinsDany_ug =`Total proteins`,
         OtherNfCN_Chla = `C and N per Chl a estimates for other Nf samples`) %>%
  # remove empty or junk columns from excel 
  select(-c(...23, ...25, ...26, ...27)) %>%
  mutate(Exp_uE = as.character(Exp_uE),
         Stress_h = as.numeric(
           recode(Stress_h, "1R" = "3"))
  )
```

```{r merge SampPrep with Catalog}
SampPrepCatalog <- left_join(x = SampPrep, y = Catalog, by = c("Sample"))
```

```{r blot files}
BlotFiles <- list.files(path = file.path(DataIn, fsep = .Platform$file.sep), pattern = ".csv", full.names = TRUE)
```

```{r Alternate read blot data, warning=FALSE, message=FALSE}
#generate read function using tidyverse read_delim
#read_delim_plus adds filename and cdate
read_delim_plus <- function(Flnm, Delimiter, HeaderRows){read_delim(file = Flnm, delim = Delimiter, skip = HeaderRows) %>%
    mutate(Filename = Flnm, cdatetime = ymd_hms(file.info(Flnm)$ctime))
}

#map_df fails converting numeric to factor?? use map then convert back to dataframe
BlotData <- BlotFiles %>% 
  map(~read_delim_plus(Flnm = ., Delimiter = Delimiter, HeaderRows = HeaderRows)) %>%
  enframe() %>%
  unnest(cols = value)
```

```{r tidy blot data, warning=FALSE, message=FALSE}
#rename columns
BlotData <- BlotData %>%
  rename(No = `No.`,
         Sample = Label,
         VolInt = `Volume (Int)`,
         AdjVolInt = `Adj. Vol. (Int)`,
         MeanBkgdInt = `Mean Bkgd. (Int)`, 
         AbsQuant_ug = `Abs. Quant. (ug)`, 
         RelQuant = `Rel. Quant.`,
         Pixels = `# of Pixels`, 
         MinValInt =  `Min. Value (Int)`, 
         MaxValInt = `Max. Value (Int)`,
         MeanValint =  `Mean Value (Int)`, 
         StdDev = `Std. Dev.`, 
         Area_mm = `Area (mm2)`,
         AbsQuant_fmol = `Abs. Quant. (fmole)`)

#remove all "," from dataframe to allow conversion from factor to character to numeric
BlotData <- BlotData %>%
  mutate_if(is.character, str_replace_all, pattern = ",", replacement = "")

#convert columns to be used fromfactors to numeric where possible; Label mixed type convert to character

#remove all columns which only contain 'na'
BlotData <- BlotData  %>%
  filter_all(any_vars(!is.na(.))) 

#extract target from filename; as.character needed to convert back from list
BlotData <- BlotData %>%
  mutate(Target = as.character(str_extract_all(Filename, paste(ProteinTargets, sep = "", collapse = '|'))))

#extract BlotId from filename; as.character needed to convert back from list
BlotData <- BlotData %>%
  mutate(BlotId = as.character(str_split_fixed(Filename, "/", n = 2)[,2])) %>%
  mutate(BlotId = as.character(str_split_fixed(BlotId, "\\.", n = 2)[,1]))
```

```{r merge data with metadata}
BlotData <- left_join(x = BlotData, y = SampPrepCatalog, by = c("Sample" = "Sample", "BlotId" = "BlotId" ))
```

```{r plot standards}
StandardsPlot <- BlotData %>%
  filter(Type == 'Standard') %>%
  ggplot() +
  geom_point(aes(y = as.numeric(Sample), x = AdjVolInt)) +
  facet_wrap(facets = vars(Filename), scales = "free") +
  theme_bw()

StandardsPlot + 
  labs(title = paste("Standards Fits", Project, sep = "_"))
```

Define fitting equations as functions.
x will be taken from 'as.numeric(Label)' when we run the fit on the BlotData filtered for Label == "Standard".
```{r equations}
# modgompertz_eqn <- function(Amax,Amu,lag,x){(Amax*(exp(-exp((Amu*exp(1))/Amax*(lag-x)+1))))}

LinEqn <- function(x, M, Intercept){((x*M)+ Intercept)
}

ExpEqn <- function(x, Mu, Intercept){(Intercept*exp(x*Mu))
}

LogEqn <- function(x, Pmax, Mu, Intercept){(Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))
}
```


Make more sophisticated with dynamic setting of starting values for curve fits based upon data sets
```{r fits starts}

# mu_start = (log(max(as.numeric(BlotData$Adj_Vol_Int[BlotData$Type == "Standard"], na.rm = TRUE))) - log(min(as.numeric(BlotData$Adj_Vol_Int[BlotData$Type == "Standard"], na.rm = TRUE))))/max(as.numeric(BlotData$Label[BlotData$Type == "Standard"], na.rm = TRUE))

LinEqnStart <- list(M = 1, Intercept = 1)
ExpEqnStart <- list(Mu = 0.0000001, ExpIntercept = 1)
LogEqnStart <- list(Pmax = 1000, LogMu = 0.00001, LogIntercept = 1)
                  
#                   
#                   logistic_eqn_start<-list(Pmax = max(fit_data[fit_data$Wavelength == nm, fitvar], na.rm = TRUE), mu = mu_start, intercept = min(fit_data[fit_data$Wavelength == nm, fitvar], na.rm = TRUE))
# 
# logistic_eqn_lower<-c((max(fit_data[fit_data$Wavelength == nm, fitvar], na.rm = TRUE) * 0.5),mu_start/10,((min(fit_data[fit_data$Wavelength == nm, fitvar], na.rm = TRUE) * 0.5)))
# 
# logistic_eqn_upper<-c((max(fit_data[fit_data$Wavelength == nm, fitvar], na.rm = TRUE) * 2),mu_start*2,((min(fit_data[fit_data$Wavelength == nm, fitvar], na.rm = TRUE) * 2)))

```

Fit linear, exponential and (later logistic) standard curves using nest purrr:map & broom::augment
Fit 'backwards' with standard fmol loads lane-1 on Y and Adj_Vol_Int to simplify later predictions of sample fmol lane-1
```{r standard fits}
BlotNest <- BlotData %>%
  filter(Type == "Standard") %>%
  mutate(Sample = as.numeric(Sample)) %>%
  nest(data = -c(Filename))

BlotFits <- BlotNest %>% 
  mutate(
  FitLin = map(data, ~nlsLM(Sample ~ LinEqn(x = AdjVolInt, M, Intercept), data = .x, start = LinEqnStart)),
    PredictLin = map(FitLin, augment),
    TidiedLin = map(FitLin, tidy),
    ParamLin = map(FitLin, glance),
  FitExp = map(data, ~nlsLM(Sample ~ ExpEqn(x = AdjVolInt, Mu, ExpIntercept), data = .x, start = ExpEqnStart)),
    PredictExp = map(FitExp, augment),
    TidiedExp = map(FitExp, tidy),
    ParamExp = map(FitExp, glance))
  
BlotPredict <- BlotFits %>%
  unnest(col = c(PredictLin, PredictExp), names_repair = "universal") %>%
  select(-c(`Sample...11`, `AdjVolInt...12`)) %>%
  rename(Sample =`Sample...4`,
         AdjVolInt = `AdjVolInt...5`,
         LinFit = `.fitted...6`, 
         ExpFit = `.fitted...13`) %>%
   select(-c(FitLin, TidiedLin, ParamLin, FitExp, TidiedExp, ParamExp))

BlotLinParam <- BlotFits %>% 
  unnest_wider(col = TidiedLin) %>%
  select(-c(data, FitLin, PredictLin, ParamLin, PredictExp, ParamExp, TidiedExp, FitExp))

BlotExpParam <- BlotFits %>% 
  unnest_wider(col = TidiedExp) %>%
  select(-c(data, FitExp, PredictExp, ParamExp, PredictExp, ParamExp, FitExp))

```


```{r plot standard fits}
StandardFitPlot <-  BlotPredict %>%
  ggplot() +  
  geom_line(aes(x = AdjVolInt, y = LinFit, colour = "Linear"), size = 0.5) +
  geom_point(aes(x = AdjVolInt, y = Sample ), size = 0.6) +
  geom_line(aes(x = AdjVolInt, y = ExpFit, colour = "Exponential", ), size = 0.5, linetype = "dashed") +
  scale_color_manual(name = "Fits", values = c("Linear" = "black", "Exponential" = "red")) +
 facet_wrap(facets = vars(Filename), scales = "free") +
  labs(y = "fmol standard lane-1") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7), axis.text.y = element_text(size = 7), strip.text.x = element_text(size = 6))
 
StandardFitPlot + labs(title = paste("Standards Fits", Project, sep = "_"))
```
 
 
```{r generate predicted fmol lane-1 values, message = FALSE, warning = FALSE}
BlotQuantData <- left_join(x = BlotData, y = BlotLinParam, by = "Filename", suffix = c("_data","_lin")) %>%
  unnest_wider(col = estimate) %>%
  rename(M = `...1`, Intercept = `...2`) %>%
  unnest_wider(col = std.error) %>%
  rename(M_sd = `...1`, Intercept_sd = `...2`) %>%
  unnest_wider(col = statistic) %>%
  rename(M_stat = `...1`, Intercept_stat = `...2`) %>%
  unnest_wider(col = p.value) %>%
  rename(M_stat_p = `...1`, Intercept_P = `...2`) %>%
  select(-c(term))

BlotQuantData <- left_join(x = BlotQuantData, y = BlotExpParam, by = "Filename") %>%
  unnest_wider(col = estimate) %>%
  rename(Mu = `...1`, ExpIntercept = `...2`) %>%
  unnest_wider(col = std.error) %>%
  rename(Mu_sd = `...1`, ExpIntercept_sd = `...2`) %>%
  unnest_wider(col = statistic) %>%
  rename(Mu_stat = `...1`, ExpIntercept_stat = `...2`) %>%
  unnest_wider(col = p.value) %>%
  rename(Mu_stat_p = `...1`, ExpIntercept_P = `...2`) %>%
  select(-c(term))

BlotQuantData <- BlotQuantData %>%
  mutate(Lin_fmol_lane = (AdjVolInt * M) + Intercept) %>%
  mutate(Exp_fmol_lane = (ExpIntercept * exp(AdjVolInt * Mu))) %>%
  select(Sample, Type, AdjVolInt, AbsQuant_fmol, Lin_fmol_lane, Exp_fmol_lane,  everything())

#Mireille modified to flag (1) fmole_lane within 10% outside standard curve, (0) if greater than 10% outside standard range..
BlotQuantData <- BlotQuantData %>%
  group_by(Filename) %>%
  mutate(
    ExpFlag = if_else(min(as.numeric(Sample[Type == "Standard"]) - min(as.numeric(Sample[Type == "Standard"]) * FlagRange)) < Exp_fmol_lane & max(as.numeric(Sample[Type == "Standard"]) * (1 + FlagRange)) > Exp_fmol_lane, 1, 0),
    LinFlag = if_else(min(as.numeric(Sample[Type == "Standard"]) - min(as.numeric(Sample[Type == "Standard"]) * FlagRange)) < Lin_fmol_lane & max(as.numeric(Sample[Type == "Standard"]) * (1 + FlagRange)) > Lin_fmol_lane, 1, 0)) %>%
  ungroup()

BlotQuantData <- BlotQuantData %>%
  select(ExpFlag, LinFlag, everything()) %>%
  select(-c(FitLin, PredictLin, TidiedLin, ParamLin)) 

#Version that works without running unnest_wider on the 'estimate' column.
#BlotQuantData <- BlotQuantData %>%
#  mutate(lin_fmol_lane = (Adj_Vol_Int * estimate[[1]][1]) + estimate[[1]][2]) %>%
#  select(Label, Type, Adj_Vol_Int, Abs_Quant_fmol, lin_fmol_lane, everything())
```

# Get Total Protein Data

```{r total protein import}
# List all files in directory
TotalProteinPath <- file.path(DataIn, TotalProtein)
TotalProteinFiles <- list.files(TotalProteinPath, full.names = TRUE)

# Separate metadata and plate reads
MetaFiles <- TotalProteinFiles[grepl("_Meta.xlsx$", TotalProteinFiles)]
PlateXMLFiles <- TotalProteinFiles[grepl(".xml$", TotalProteinFiles)]

# import all plate reads
PlateRead <- NULL

# loop through all plate xml files
for(i in PlateXMLFiles) {
  # use softmax library to read xml
  XMLData.i <- read_softmax_xml(i)
  # convert to data frame
  PlateRead.i <- as.data.frame(XMLData.i)
  # append to PlateRead df
  PlateRead <- bind_rows(PlateRead, PlateRead.i)
}

# import all metadata
Meta <- NULL
# loop through meta files
for(MetaFile in MetaFiles){
  MetaSheets <- readxl::excel_sheets(MetaFile)

  # loop through excel sheets
  for(sheet in MetaSheets){
    # Read first line to get plate name
    MetaPlateName <- names(readxl::read_xlsx(MetaFile, sheet = sheet, n_max = 1))[1] 
    # read data from below 1st line
    Data <- readxl::read_xlsx(MetaFile, sheet = sheet, skip = 1) %>%
    # convert all to character
    mutate_all(as.character) %>%
    # wide to long data
    pivot_longer(-Row, names_to = "Column", values_to = "SampleName") %>%
    mutate(Well = paste(Row, Column, sep = ""),
           PlateBaseName = MetaPlateName) %>%
      filter(!is.na(SampleName))
    print(MetaPlateName)
    # append current sheet data to Meta df
    Meta <- bind_rows(Meta, Data)
  }
}
```

```{r subtract blanks}
BlankSubtracted <- PlateRead %>%
  # If plate name starts with blank, classify as Blank, otherwise classify as Sample
  mutate(WellType = if_else(grepl("^Blank", Plate), "Blank", "Sample"),
         # Get base name by removing Blank from plate name (if present)
         PlateBaseName = gsub("^Blank", "", Plate)) %>%
  select("Wavelength", "WellType", "Well", "Value", "PlateBaseName") %>%
  # convert from long to wide format
  pivot_wider(names_from = WellType, values_from = Value, names_prefix = "Value") %>%
  # subtract blank from each sample
  mutate(BlankSubtractedValue = ValueSample - ValueBlank)
```

```{r join plate values with metadata}
JoinedProtein <- left_join(Meta, BlankSubtracted, by = c("Well", "PlateBaseName")) %>%
  # If sample starts with "P", classify as Unknown, otherwise classify as Standard
  mutate(SampleType = if_else(grepl("^P", SampleName), "Unknown", "Standard"))

JoinedProtein %>%
  arrange(SampleName)
```

```{r functions for mapping protein standards}
FilterStds <- function(Data) {
  Data %>%
    filter(SampleType == "Standard") %>%
    mutate(SampleName = as.numeric(SampleName))
}

FitStdCurves <- function(Data) {
  lm(SampleName ~ BlankSubtractedValue, data = Data)
}

GetRsq <- function(Model) {
  broom::glance(Model)$r.squared
}
```

```{r nest and add standard curve lms}
Nested <- JoinedProtein %>%
  # Separate each plate into its own dataframe
  group_by(PlateBaseName) %>%
  nest() %>%
  # isolate standards in own dataframe
  mutate(StdData = map(data, FilterStds)) %>%
  # apply linear model to each standard dataframe
  mutate(Model = map(StdData, FitStdCurves)) %>%
  # get r^2 for each model
  mutate(Rsq = map(Model, GetRsq)) %>%
  mutate_at("Rsq", unlist)
```

```{r protein conc from absorbance and standard curve models}
ProteinData <- NULL

# loop through each plate (each row of nested dataframe)
for(i in 1:nrow(Nested)) {
  # get data and model from current row, need to extract with [[]] since in a list
  Data <- Nested$data[i][[1]]
  Model <- Nested$Model[i][[1]]
  
  PlateBaseName.i <- Nested$PlateBaseName[i]
  
  # use model to predict protein conc. from absorbanc value and corresponding standard curve
  FittedData <- Data %>%
    modelr::add_predictions(Model) %>%
     mutate(PlateBaseName = PlateBaseName.i)
  
  ProteinData <- bind_rows(ProteinData, FittedData)
}
ProteinData
# drop all negative predicted values 
# use ifelse not if_else since NA and double are different types 
  
  

# keep only useful info from plate data
LongProteinData <- ProteinData %>%
  # only keep samples (starts with P)
  filter(grepl("^P", SampleName)) %>%
  # check if all 3 replicates are negative
  group_by(SampleName) %>%
  mutate(AllNegativeProt = (3 == sum((pred < 0)))) %>%
  # set replicates as NA if negative
  mutate(pred = ifelse(pred < 0, NA, pred)) %>%
  # subset columns
  select(SampleName, "Protein_ug_uL" = pred, PlateBaseName, AllNegativeProt) %>%
  # calculate mean for each sample
  group_by(SampleName) %>%
  mutate(AvgTotalProt = mean(Protein_ug_uL,
                             na.rm = TRUE),
         StdDev = sd(Protein_ug_uL,
                     na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(SampleName)

LongProteinData
```

```{r long to wide}
# pivot long data to wide, column for each replicate
WideProteinData <- LongProteinData %>%
  # rename duplicate entries, append .# to end
  mutate(across(SampleName, make.unique, sep = "___")) %>%
  # separate replicate # from sample name
  separate(SampleName, c("SampleName", "Replicate"), sep = "___") %>%
  # make.unique starts counting at 1, first entry is untouched.
  # so replicate column for first entry will be "NA". Replace with 0.
  mutate(Replicate = if_else(is.na(Replicate), "0", Replicate)) %>%
  # pivot to wide format, column for each replicate.
  pivot_wider(names_from = Replicate, values_from = Protein_ug_uL) %>%
  mutate(CV = StdDev / AvgTotalProt * 100) %>%
  # rename and select columns
  select(SampleName, TotalProtRep1 = "0", 
         TotalProtRep2 = "1", TotalProtRep3 = "2",
         AvgTotalProt, CV, StdDev, PlateBaseName, AllNegativeProt)

WideProteinData
```

```{r join protein data with BlotQuantData}
BlotQuantData <- BlotQuantData %>%
  # join imported total protein data with BlotQuantData
  left_join(., 
          WideProteinData, 
          by = c("Sample" = "SampleName" )) %>%
  # rename protein columns to fit with later format
  rename(Rep1TotalProteinExtract_ug_uL = TotalProtRep1,
         Rep2TotalProteinExtract_ug_uL = TotalProtRep2,
         Rep3TotalProteinExtract_ug_uL = TotalProtRep3,
         AverageTotalProteinExtract_ug_uL = AvgTotalProt,
         TotalProteinExtractFileName = PlateBaseName) %>%
  # re-parse to read as numeric
  type_convert() %>%
  arrange(Sample) %>%
  mutate(TotalProteinLoaded_ug = AverageTotalProteinExtract_ug_uL * ExtractLoaded_uL,
         TotalProteinExtract_ug = AverageTotalProteinExtract_ug_uL * ExtractVolume_uL,
         TotalChlaExtract_ug = TotalChla_ug)
```


# Plot quantitation

Plot the quantitation points overlaid on the standards plots
```{r quant_plot}
QuantLinPlot <- StandardFitPlot +
  geom_point(data = BlotQuantData, aes(x = AdjVolInt, y = Lin_fmol_lane, colour = as.factor(LinFlag)), size = 0.2) +
  scale_color_manual(values = c("orange","green","black", "grey")) +
  labs(title = "Quantitation Linear Fits")

QuantLinPlot

QuantExpPlot <- StandardFitPlot +
  geom_point(data = BlotQuantData, aes(x = AdjVolInt, y = Exp_fmol_lane, colour = as.factor(ExpFlag)), size = 0.2) +
  scale_color_manual(values = c("orange","green","black", "grey")) +
  labs(title = "Quantitation Exponential Fits")

QuantExpPlot
```

```{r fmol target per mL sample}
BlotQuantData <- BlotQuantData %>%
         mutate(Lin_fmol_mL = Lin_fmol_lane * (as.numeric(ExtractVolume_uL)/as.numeric(ExtractLoaded_uL)) * (1/`Harvest_mL`), na.rm = TRUE,
           Exp_fmol_mL = Exp_fmol_lane *  (as.numeric(ExtractVolume_uL)/as.numeric(ExtractLoaded_uL)) * (1/`Harvest_mL`), na.rm = TRUE)
```

```{r fmol target per ug Chla}
BlotQuantData <- BlotQuantData %>%
  mutate(ChlaLoaded_ug = TotalChla_ug / ExtractVolume_uL * ExtractLoaded_uL,
    Lin_fmol_ugChla = Lin_fmol_lane * 1/as.numeric(ChlaLoaded_ug), na.rm = TRUE,
         Exp_fmol_ugChla = Exp_fmol_lane * 1/as.numeric(ChlaLoaded_ug), na.rm = TRUE)
```

Mireille added this
```{r fmol target per ug total protein}
BlotQuantData <- BlotQuantData %>%
  mutate(Lin_fmol_ugTotalProtein = Lin_fmol_lane * 1/as.numeric(TotalProteinLoaded_ug), na.rm = TRUE,
         Exp_fmol_ugTotalProtein = Exp_fmol_lane * 1/as.numeric(TotalProteinLoaded_ug), na.rm = TRUE)

```

Mireille added this
```{r Total protein per mL harvested}
BlotQuantData <- BlotQuantData %>%
mutate(TotalProtein_ug_mLHarvest = as.numeric(AverageTotalProteinExtract_ug_uL) * as.numeric(ExtractVolume_uL)* 1/(`Harvest_mL`), na.rm = TRUE)
```

```{r save BlotQuantData}
saveRDS(BlotQuantData, file = file.path(DataOut,paste(Project, "BlotQuantData", ".Rds",sep = "_"),fsep = .Platform$file.sep))

write_csv(BlotQuantData, path = file.path(DataOut,paste(Project, "_BlotQuantData",".csv",sep = ""),fsep = .Platform$file.sep))

ggsave(filename = file.path(DataOut, paste(Project, "QuantLinPlot.png", sep = "_"),fsep = .Platform$file.sep), device = "png", plot = QuantLinPlot)

ggsave(filename = file.path(DataOut, paste(Project, "QuantExpPlot.png", sep = "_"),fsep = .Platform$file.sep), device = "png", plot = QuantExpPlot)
```

```{r compile a condensed report}
#Filter Samples starting with "P[1-9]" (all samples beginning with "P" followed immediately by a numeral).
# use M_stat_p and Mu_stat_p to determine which fit, linear or exponential to report 

TargetProteinReport <- BlotQuantData %>%
  filter(Type == "Unknown") %>%
  filter(grepl("^P[0-9]{1}", Sample)) %>%
  mutate(fmol_mL = if_else(((M_stat_p) < (Mu_stat_p)), Lin_fmol_mL, Exp_fmol_mL)) %>%
  mutate(fmol_ugChla = if_else(((M_stat_p) < (Mu_stat_p)), Lin_fmol_ugChla, Exp_fmol_ugChla)) %>%
  mutate(fmol_ugTotalProtein = if_else(((M_stat_p) < (Mu_stat_p)), Lin_fmol_ugTotalProtein,   Exp_fmol_ugTotalProtein)) %>%
  mutate(QuantModel = if_else(((M_stat_p) < (Mu_stat_p)), "Linear", "Exponential")) %>%
  mutate(Flag = if_else(((M_stat_p) < (Mu_stat_p)), LinFlag, ExpFlag)) %>%
  mutate(ProtChlorRatio = TotalProteinExtract_ug / TotalChlaExtract_ug) %>%
  select(Sample, 
         Date,
         Sp,
         Exp, 
         CultureId, 
         Growth_uE, 
         Exp_uE, 
         Stress_h,
         Inhibitor,
         DarkPre_h,
         Harvest_mL,
         Chla_ug_L,
         TotalChla_ug,
         Carbon_ug_mL,
         TotalCarbon_ug,
         N_ug_mL,
         TotalN_ug,
         ExtractVolume_uL,
         BlotId,
         Target,
         BlotLane,
         ExtractLoaded_uL,
         Rep1TotalProteinExtract_ug_uL,
         Rep2TotalProteinExtract_ug_uL,
         Rep3TotalProteinExtract_ug_uL,
         AverageTotalProteinExtract_ug_uL,
        TotalProteinExtractFileName,
         TotalProtein_ug_mLHarvest,
        TotalChlaSheetName,
         fmol_mL,
         fmol_ugChla,
         fmol_ugTotalProtein,
         QuantModel,
        Flag,
        AllNegativeProt,
        SamplePrepComments,
        TotalProteinExtract_ug,
        TotalChlaExtract_ug,
        ProtChlorRatio,
        BlotComments)
```

```{r saves TargetProteinReport}

saveRDS(TargetProteinReport, file = file.path(DataOut,paste(Project, "TargetProteinReport", ".Rds",sep = "_"),fsep = .Platform$file.sep))

# also save RDS file to shiny folder
saveRDS(TargetProteinReport, file = file.path("ShinyGreenEdge", "data", paste(Project, "TargetProteinReport", ".Rds",sep = "_"),fsep = .Platform$file.sep))

write_csv(TargetProteinReport, path = file.path(DataOut,paste(Project, "_TargetProteinReport",".csv",sep = ""),fsep = .Platform$file.sep))
```


```{r plot protein quant vs. growth light}
# List target proteins, which conditions to plot, protein normalize methods
Targets <- ProteinTargets
Condition <- "Growth_uE"
NormalizeMethods <- c("fmol_mL", "fmol_ugChla", "fmol_ugTotalProtein")

# initialize lists, counter
PlotList <- list()
DataList <- list()
PlotCount <- 1

# loop through target proteins
for(i in 1:length(Targets)){
  # select corresponding condition
  ConditionSym <- sym(Condition)
  
  # format/select data for plotting
  Data <- TargetProteinReport %>%
    #mutate(Stress_h = factor(Stress_h, levels = c("0", "0.5", "2", "1R"))) %>%
    filter(Target == Targets[i],
           TotalChlaSheetName == "Nf_Growthvs.IceCamp",
           !is.na(!!ConditionSym))
  # Add to list of plotting data
  DataList[[i]] <- Data
  
  # loop through methods of normalizing protein amount
  for(j in 1:length(NormalizeMethods)){
    NormalizeMethod <- sym(NormalizeMethods[j])
    Plot <- ggplot(Data) +
  geom_point(aes(x= !!ConditionSym, y= !!NormalizeMethod, 
                 colour = as.factor(Flag), shape = as.factor(BlotId))) +
  # facet wrap if stress_h is on x axis
  #{if(Conditions[i] == "Stress_h")facet_wrap(vars(CultureId))}+
  theme_bw() +
  labs(title = paste("Reported", Targets[i], NormalizeMethods[j], "harvested"))                       
  print(Plot)
  
  # add plot to list, with informative name
  PlotList[[PlotCount]] <- Plot
  names(PlotList)[PlotCount] <- paste(Targets[i], Condition,
                                            NormalizeMethods[j], sep = "_")
  # advance plot counter
  PlotCount <- PlotCount + 1
  }
}
```

```{r save exploratory growth plots}
for(i in 1:length(PlotList)){
  PlotName <- names(PlotList)[i]
  Plot <- PlotList[[i]]
  
  print(PlotName)
  ggsave(filename = file.path(PlotsPath, paste(Project, "Explore", names(PlotList)[i], ".png", sep = ""), fsep = .Platform$file.sep), device = "png", plot = PlotList[[i]])
}
```




MIrielle removed shape from wrap as there are too many now
```{r plot protein quant vs. experimental conditions}
# List target proteins, which conditions to plot, protein normalize methods
Targets <- ProteinTargets
Conditions <- c("Stress_h", "Stress_h", "Stress_h")
NormalizeMethods <- c("fmol_mL", "fmol_ugChla", "fmol_ugTotalProtein", "TotalProteinExtract_ug", "TotalChlaExtract_ug", "ProtChlorRatio")


# initialize lists, counter
PlotList <- list()
DataList <- list()
PlotCount <- 1

# loop through target proteins
for(i in 1:length(Targets)){
  # select corresponding condition
  Condition <- sym(Conditions[i])
  
  # format/select data for plotting
  Data <- TargetProteinReport %>%
    #mutate(Stress_h = factor(Stress_h, levels = c("0", "0.5", "2", "1R"))) %>%
    filter(Target == Targets[i],
           !is.na(!!Condition))
  # Add to list of plotting data
  DataList[[i]] <- Data
  
  # loop through methods of normalizing protein amount
  for(j in 1:length(NormalizeMethods)){
    NormalizeMethod <- sym(NormalizeMethods[j])
    Plot <- ggplot(Data) +
      
      geom_point(aes(x= !!Condition, y= !!NormalizeMethod, 
                 colour = Inhibitor)) +
      geom_vline(xintercept = 2, color = "red", size = 0.25) +
  # facet wrap if stress_h is on x axis
      {if(Conditions[i] == "Stress_h")facet_wrap(vars(CultureId))}+
      theme_bw() +
      labs(title = paste("Reported", Targets[i], 
                         NormalizeMethods[j], "harvested"))
      print(Plot)
  
  # add plot to list, with informative name
  PlotList[[PlotCount]] <- Plot
  names(PlotList)[PlotCount] <- paste(Targets[i], Conditions[i],
                                            NormalizeMethods[j], sep = "_")
  # advance plot counter
  PlotCount <- PlotCount + 1
  }
}
```

```{r save light stress plots}
for(i in 1:length(PlotList)){
  PlotName <- names(PlotList)[i]
  Plot <- PlotList[[i]]
  
  print(PlotName)
  ggsave(filename = file.path(PlotsPath, paste(Project, names(PlotList)[i], ".png", sep = ""), fsep = .Platform$file.sep), device = "png", plot = PlotList[[i]])
}
```


```{r}
SpeciesData <- TargetProteinReport %>%
  filter(!is.na("Stress_h"),
         Target == "PsbA",
         Sp == "Chaetoceros neogracilis") %>%
  group_by(Stress_h, Inhibitor) %>%
  mutate(Meanfmol_ugTotalProtein = mean(fmol_ugTotalProtein))

SpeciesPlot <- ggplot(SpeciesData) +
  geom_point(aes(x = Stress_h, y = Meanfmol_ugTotalProtein))+#, colour = as.factor(CultureId))) +
  facet_grid(vars(Sp), vars(Inhibitor))+
  theme_bw()+
  ggtitle("PsbA")

SpeciesPlot

ggsave("Plots/SpeciesPlot.png")
```

```{r}
Fcp6SpeciesData <- TargetProteinReport %>%
  filter(!is.na("Stress_h"),
         Target == "Fcp6",
         Sp == "Chaetoceros neogracilis") %>%
  group_by(Stress_h, Sp, Inhibitor) %>%
  mutate(Meanfmol_ugTotalProtein = mean(fmol_ugTotalProtein))

Fcp6SpeciesPlot <- ggplot(SpeciesData) +
  geom_point(aes(x = Stress_h, y = Meanfmol_ugTotalProtein))+#, colour = as.factor(CultureId))) +
  facet_grid(vars(Sp), vars(Inhibitor))+
  theme_bw()+
  ggtitle("Fcp6")

Fcp6SpeciesPlot

ggsave("Plots/Fcp6SpeciesPlot.png")
```


```{r}
Data <- TargetProteinReport %>%
  filter(!is.na("Stress_h"),
         Target == "PsbA",
         Exp != "Nf_Growth_vs.IceCamp",
         Sp == "Fragilariopsis cylindrus")
         #!grepl("Natalie$", BlotId))
        # BlotId != "20200612Fcp")	

LinePlot <- ggplot(Data) +
  geom_point(aes(x = TotalChlaExtract_ug, y = TotalProteinExtract_ug, colour = BlotId, shape = BlotId)) +
  stat_smooth(aes(x = TotalChlaExtract_ug, y = TotalProteinExtract_ug),   method = "lm", fullrange = TRUE)+
  coord_cartesian(xlim = c(0,10), ylim = c(0, 150))+
  theme_bw() 

LinePlot

ggsave("Plots/LinePlot.png")
  
```

Next want to do the same light stress plots, but for each species.
This could be done the same way with another nested for loop (3 total), but hard to read.  
Instead, create a dataframe of all 8 (`2*1*2*2`) possible combinations with `expand.grid()`.  

```{r}
Targets <- c("Fcp6", "PsbA")
Conditions <- c("Stress_h")
NormalizeMethods <- c("fmol_mL", "fmol_ugChla", "fmol_ugTotalProtein")
Species <- c("Nitszchia frigida", "Fragilariopsis cylindrus")
Test <- c(1, 2, 3)

CombinationMatrix <- expand.grid("Targets" = Targets, 
                                 "Conditions" = Conditions,
                                 "NormalizeMethods" = NormalizeMethods, 
                                 "Species" = Species)
CombinationMatrix
```

Combine all into a single for loop that goes row-by-row through combination matrix.

```{r plot protein quant vs. experimental conditions-plot for each species}
# initialize lists, counter
  PlotList <- list()
  DataList <- list()
  PlotCount <- 1


for(i in 1:nrow(CombinationMatrix)) {
  Targets.i <- as.character(CombinationMatrix$Targets[i])
  Conditions.i <- as.character(CombinationMatrix$Condition[i])
  NormalizeMethods.i <- as.character(CombinationMatrix$NormalizeMethods[i])
  Species.i <- as.character(CombinationMatrix$Species[i])
  
  
  
  # format/select data for plotting
  Data <- TargetProteinReport %>%
    #mutate(Stress_h = factor(Stress_h, levels = c("0", "0.5", "2", "1R"))) %>%
    filter(Target == Targets.i,
           TotalChlaSheetName == "LightStress_IceCamp_mimic",
           Sp == Species.i,
           !is.na(!!Conditions.i))
  # Add to list of plotting data
  DataList[[i]] <- Data
  
  # loop through methods of normalizing protein amount
  NormalizeMethod <- sym(NormalizeMethods.i)
  Plot <- ggplot(Data) +
    geom_point(aes(x= !!Condition, y= !!NormalizeMethod, 
               shape = as.factor(Flag), colour = as.factor(CultureId))) +
    geom_vline(xintercept = 2, color = "red", size = 0.25) +
    # facet wrap if stress_h is on x axis
    {if(Conditions.i == "Stress_h")facet_wrap(vars(Inhibitor))}+
    theme_bw() +
    labs(title = paste("Reported", Targets.i, NormalizeMethods.i, "harvested", Species.i))                       
  print(Plot)
  
  # add plot to list, with informative name
  PlotList[[PlotCount]] <- Plot
  names(PlotList)[PlotCount] <- paste(Targets.i, 
                                      Conditions.i,
                                      NormalizeMethods.i, 
                                      Species.i,
                                      sep = "_")
  # advance plot counter
  PlotCount <- PlotCount + 1
}

```

```{r save plots for each species}
for(i in 1:length(PlotList)){
  PlotName <- names(PlotList)[i] %>%
    # replace long and space-containing species names with abbreviations
    gsub("Nitszchia frigida", "Nf", .) %>%
    gsub("Fragilariopsis cylindrus", "Fc", .)
  
  Plot <- PlotList[[i]]
  
  print(PlotName)
  ggsave(filename = file.path(PlotsPath, paste(Project, names(PlotList)[i], ".png", sep = ""), fsep = .Platform$file.sep), device = "png", plot = PlotList[[i]])
}
```
